---
title: "Gapminder Analysis"
author: "Patrice Crosby"
date: "2024-02-07"
output: 
    html_document:
        toc: yes
        toc_float: yes
        code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the first line of my document. In this document, I will analyze how life expectancy has changed over time.

# reading in gapminder data

```{r reading-in-data}
gapminder_df <- read.csv("data/gapminder_data.csv")
```

## smaller header

### an even smaller header

#### a very small header

Lists:

*list item 1* list item 2

Another type of list:

-Hyphen list -Hyphen list

Numbered lists:

1.  Number one
2.  Number two

-   Sub list

3.  Number three

We can make text **bold with two asterisks** or *italic with one asterisk*

We can also insert greek letters $\mu$ or $\Delta$

We can make [hyperlinks](https://pad.carpentries.org/Spring2024-BIOMI6300)

#Calculating summary statistics

```{r calculating-median-pop, echo=FALSE}
med_pop <- median(gapminder_df$pop)

num_obs <- nrow(gapminder_df)
```

Over `r num_obs` observations, the median population was `r round(med_pop/10e6, 3)` million people.

\# Plotting life expectancy over the years

```{r line graph, eval=FALSE}
library(ggplot2)

ggplot(gapminder_df, aes(x= year, y=lifeExp, group=country, color=continent))+
    geom_line()+
    scale_y_log10() +
    scale_x_continuous(limits = c(1997,2007))

```

# Plotting Population Distribution

```{r histogram, message=FALSE, fig.width=10}

ggplot(gapminder_df, aes(x = pop)) +
    geom_histogram()
```

```{r}

```
